---
title: "Demographics Correction Script"
author: "OPMI"
date: "1/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(magrittr)

# avoid scientific notation
options(scipen = 999)
```

Purpose of the script is to streamline the demographic correction process -- i.e., take StreetLight home and work analyses and create correction factors for origins in order to bring O-D analysis demographics into agreement with Title VI definitions of racial/ethnic minority & low-income. In addition, the script can add information such as limited English proficiency. 

When running for a new time period (currently set up to run Fall 2019), be sure to run all chunks currently marked as `eval = FALSE`. To save processing time, the script skips these chunks and loads from RData files stored in Data/.

# Change File Paths

```{r}
# List the file paths for your username and StreetLight analyses below
user <- "path_to_username"

# # Fall 2021
# za <- "364709_Fall_2021_Home_Grids"
# bike <- "364336_Copy_MBTA_Bike_Fall_2021_Hourly"
# ped <- "364337_Copy_MBTA_Ped_Fall_2021_Hourly_Demos"
# veh <- "364293_MBTA_Service_Area_Fall_2021_Hourly"
# veh_trips_col_types = "ccccdccccdcccddddddddddddddddddddddddddddddddddddddddddddddddddd"
# 
# od_output_name <- "Corrected_Fall_2021_OD_Volumes"


# 2019
za <- "202270_Pre_Pandemic_Home_Grids_2019_MBTA"
bike <- "200555_MBTA_Service_Area_Bike_Fall_2019"
ped <- "200573_MBTA_Service_Area_Ped_Fall_2019"
veh <- "200524_MBTA_Service_Area_Fall_2019_Hourly"
veh_trips_col_types <- "ccccdccccdcccdddddddddddddddddddddddddddddddddddddddddd"

od_output_name <- "Corrected_Fall_2019_OD_Volumes"
```

# Streetlight Names

```{r}
stl_names <- read_csv("./Data/StreetLight_ColNames.csv", col_types = "cc") %>% deframe()
```


# BAZ Reference

```{r}
baz_ref <- read_csv(paste0(user, "filepath"),
                    col_types = "ccddcdccdcd")
baz_ref %<>% select(baz, baz_name, blkgrp)
```



# Census

```{r, eval = FALSE}
# load census key with census_api_key in console

# load variable crosswalk 
# vars_2019_acs <- load_variables(2019, "acs5")
# save(vars_2019_acs, file = "./Data/vars_2019_acs.RData")
load("./Data/vars_2019_acs.RData")

# load MA BG variables
census_mass_bgs_vars <-
  c("B03002_001","B03002_002","B03002_003","B03002_012","B03002_013",
    "B19001_001", "B19001_002", "B19001_003", "B19001_004", "B19001_005", "B19001_006",
    "B19001_007", "B19001_008", "B19001_009", "B19001_010",
    "C16002_001", "C16002_004", "C16002_007", "C16002_010", "C16002_013")

# pull variables 
census_mass_bgs <-
  get_acs(geography = "block group",
          variables = census_mass_bgs_vars,
          output = "tidy",
          state = "MA",
          year = 2019,
          show_call = TRUE)
Sys.sleep(1)


# MA tract variables
census_mass_cts_vars <- 
  c("B03002_001", 
    "B08201_001", "B08201_008", "B08201_014", "B08201_015", "B08201_020",
    "B08201_021", "B08201_026", "B08201_027", "B08201_028")

# pull variables 
census_mass_cts <-
  get_acs(geography = "tract",
          variables = census_mass_cts_vars,
          output = "tidy",
          state = "MA",
          year = 2019,
          show_call = TRUE)
Sys.sleep(1)


# variables for states
census_states_vars <- c(census_mass_bgs_vars, census_mass_cts_vars) %>% unique()
rm(census_mass_bgs_vars, census_mass_cts_vars)

# variables crosswalk
census_vars_xwalk <-
  tibble(acs_var = census_states_vars) %>%
  left_join(vars_2019_acs, by = c("acs_var" = "name")) %>%
  mutate(label = str_remove(label, "Estimate!!Total:"), label = str_remove_all(label, "!!"), 
         label = case_when(acs_var == "B03002_001" ~ "Total Population",
                           acs_var %in% c("B19001_001", "C16002_001", "B08201_001") ~ "Total HH", TRUE ~ label))
rm(vars_2019_acs)


# pull state data
census_states <-
  get_acs(geography = "state",
          variables = census_states_vars,
          output = "tidy",
          year = 2019,
          show_call = TRUE)
rm(census_states_vars)



## Create Demographics Tables


### Pivot 

census_vars_xwalk %<>%
  mutate(colname = recode(label,
         "Total Population" = "pop_tot",
         "Not Hispanic or Latino:" = "pop_nothisp",
         "Not Hispanic or Latino:White alone" = "pop_nothisp_white",
         "Hispanic or Latino:" = "pop_hisp",
         "Hispanic or Latino:White alone" = "pop_hisp_white",
         "Total HH" = "hh_tot", 
         "Less than $10,000" = "hh_lessthan10k",
         "$10,000 to $14,999" = "hh_10to15k",
         "$15,000 to $19,999" = "hh_15to19k",
         "$20,000 to $24,999" = "hh_20to25k",
         "$25,000 to $29,999" = "hh_25to30k",
         "$30,000 to $34,999"= "hh_30to35k",
         "$35,000 to $39,999" = "hh_35to40k",
         "$40,000 to $44,999" = "hh_40to45k",
         "$45,000 to $49,999"  = "hh_45to50k",
         "Total HH" = "hh_tot",
         "Spanish:Limited English speaking household" = "hh_lep_spanish",
         "Other Indo-European languages:Limited English speaking household"  = "hh_lep_indoeur",
         "Asian and Pacific Island languages:Limited English speaking household" = "hh_lep_asian",
         "Other languages:Limited English speaking household" = "hh_lep_other",
         "Total HH" = "hh_tot",
         "1-person household:No vehicle available" = "hh_1pers_nocar",
         "2-person household:No vehicle available" = "hh_2pers_nocar",
         "2-person household:1 vehicle available" = "hh_2pers_1car",
         "3-person household:No vehicle available" = "hh_3pers_nocar",
         "3-person household:1 vehicle available"  = "hh_3pers_1car",
         "4-or-more-person household:No vehicle available" = "hh_4pers_nocar",
         "4-or-more-person household:1 vehicle available" = "hh_4pers_1car",
         "4-or-more-person household:2 vehicles available" = "hh_4pers_2car"
  ),
  .after = label) 

census_mass_bgs_wide <-
  census_mass_bgs %>% left_join((census_vars_xwalk %>% select(acs_var, colname)), by = c("variable" = "acs_var")) %>% 
  select(GEOID, colname, estimate) %>% 
  distinct() %>% 
  pivot_wider(names_from = colname, values_from = estimate)
rm(census_mass_bgs)

census_mass_cts_wide <-
  census_mass_cts %>% left_join((census_vars_xwalk %>% select(acs_var, colname)), by = c("variable" = "acs_var")) %>% 
  select(GEOID, colname, estimate) %>% 
  distinct() %>% 
  pivot_wider(names_from = colname, values_from = estimate)
rm(census_mass_cts)

census_states_wide <-
  census_states %>% left_join((census_vars_xwalk %>% select(acs_var, colname)), by = c("variable" = "acs_var")) %>% 
  select(GEOID, NAME, colname, estimate) %>% 
  distinct() %>% 
  pivot_wider(names_from = colname, values_from = estimate)
rm(census_states)

rm(census_vars_xwalk)



### Join tracts to BGs 


census_mass_cts_pcts <-
  census_mass_cts_wide %>% 
  mutate(hh_limveh_strict = 
           hh_1pers_nocar + hh_2pers_nocar + hh_3pers_nocar + hh_3pers_1car + hh_4pers_nocar + hh_4pers_1car,
         hh_limveh_weak = 
           hh_1pers_nocar + hh_2pers_nocar + hh_2pers_1car + hh_3pers_nocar + hh_3pers_1car + hh_4pers_nocar + hh_4pers_1car + hh_4pers_2car
           ) %>% 
  mutate(hh_limveh_strict_pct = hh_limveh_strict/hh_tot, hh_limveh_weak_pct = hh_limveh_weak/hh_tot) %>% 
  select(GEOID, ends_with("pct"))
rm(census_mass_cts_wide)

census_mass_bgs_pcts <-
  census_mass_bgs_wide %>% 
  mutate(
    pop_white = pop_nothisp_white + pop_hisp_white,
    pop_mino = pop_tot - pop_nothisp_white,
    hh_li45 = hh_lessthan10k + hh_10to15k + hh_15to19k + hh_20to25k + hh_25to30k + hh_30to35k + hh_35to40k + hh_40to45k,
    hh_li50 = hh_li45 + hh_45to50k,
    hh_lep = hh_lep_spanish + hh_lep_indoeur + hh_lep_asian + hh_lep_other) %>% 
  mutate(
    pop_white_pct = pop_white/pop_tot,
    pop_hisp_white_pct = pop_hisp_white/pop_tot,
    pop_mino_pct = pop_mino/pop_tot,
    pop_hisp_pct = pop_hisp/pop_tot,
    hh_li45_pct = hh_li45/hh_tot,
    hh_li50_pct = hh_li50/hh_tot,
    hh_lep_pct = hh_lep/hh_tot) %>% 
  select(GEOID, pop_tot, hh_tot, starts_with("pop_mino"), starts_with("pop_white"), starts_with("pop_hisp"), starts_with("hh_li"), hh_lep, hh_lep_pct)
rm(census_mass_bgs_wide)

census_mass_info <-
  census_mass_bgs_pcts %>% 
  mutate(tosep = GEOID, .after = GEOID) %>% 
  separate(tosep, sep = 11, into = c("tract", "bg")) %>% 
  left_join(census_mass_cts_pcts, by = c("tract" = "GEOID")) %>% 
  select(-c(tract, bg))
rm(census_mass_bgs_pcts, census_mass_cts_pcts)




### Clean States table


census_states_info <-
  census_states_wide  %>% 
  mutate(hh_limveh_strict = 
           hh_1pers_nocar + hh_2pers_nocar + hh_3pers_nocar + hh_3pers_1car + hh_4pers_nocar + hh_4pers_1car,
         hh_limveh_weak = 
           hh_1pers_nocar + hh_2pers_nocar + hh_2pers_1car + hh_3pers_nocar + hh_3pers_1car + hh_4pers_nocar + hh_4pers_1car + hh_4pers_2car
           ) %>% 
  mutate(hh_limveh_strict_pct = hh_limveh_strict/hh_tot, hh_limveh_weak_pct = hh_limveh_weak/hh_tot) %>% 
  mutate(
    pop_white = pop_nothisp_white + pop_hisp_white,
    pop_mino = pop_tot - pop_nothisp_white,
    hh_li45 = hh_lessthan10k + hh_10to15k + hh_15to19k + hh_20to25k + hh_25to30k + hh_30to35k + hh_35to40k + hh_40to45k,
    hh_li50 = hh_li45 + hh_45to50k,
    hh_lep = hh_lep_spanish + hh_lep_indoeur + hh_lep_asian + hh_lep_other) %>% 
  mutate(
    pop_white_pct = pop_white/pop_tot,
    pop_hisp_white_pct = pop_hisp_white/pop_tot,
    pop_hisp_pct = pop_hisp/pop_tot,
    pop_mino_pct = pop_mino/pop_tot,
    hh_li45_pct = hh_li45/hh_tot,
    hh_li50_pct = hh_li50/hh_tot,
    hh_lep_pct = hh_lep/hh_tot) %>% 
  select(GEOID, NAME, pop_tot, hh_tot, starts_with("pop_mino"), starts_with("pop_white"), starts_with("pop_hisp"),
         hh_li45, hh_li50, hh_li45_pct, hh_li50_pct, hh_lep, hh_lep_pct, hh_limveh_strict, hh_limveh_strict_pct, 
         hh_limveh_weak, hh_limveh_weak_pct)
rm(census_states_wide)

save(census_mass_info, census_states_info, file = "./Data/Census_Data.RData")
```

```{r}
load("./Data/Census_Data.RData")
```


# Residents Table

```{r}
# join census pop totals on block group
baz_pops <-
  left_join(baz_ref, 
          (census_mass_info %>% select(GEOID, pop_tot)),
          by = c("blkgrp" = "GEOID")) %>% 
  rename(bg_pop = pop_tot)

# calculate percent of baz by block group
baz_pops %<>% 
  group_by(baz, baz_name) %>% 
  mutate(baz_pop = sum(bg_pop)) %>% 
  ungroup %>% 
  mutate(bg_pct_baz = bg_pop/baz_pop)
```



# Streetlight HWL

## BGs

```{r, eval = FALSE}
# read home block groups
za_hbg <- read_csv(paste0("./Data/", za, "/Home Work/", za, "_home_block_groups_all.csv"),
                   col_types = "ccccccccdd",
                   na = c("", "N/A"))
# rename
za_hbg <- rename(za_hbg, any_of(stl_names))

# filter to all days, all hours (for highest sample size), massachusetts block groups
za_hbg_filtered <-
  filter(za_hbg,
         day_type == "0: All Days (M-Su)" &
         day_part == "0: All Day (12am-12am)" &
         str_detect(geoid, "^25"))
rm(za_hbg)
gc()

# write to RData file
save(za_hbg_filtered, file = paste0("./Data/", za, "_hbg_filtered.RData"))
```

```{r}
load(paste0("./Data/", za, "_hbg_filtered.RData"))
```

HWL analysis should contain visitors and/or workers for block groups that are not in the Zone of Analysis, and only workers for block groups that are contained within the Zone of Analysis. 
  
```{r}
# tag block groups as being within the zone of analysis or not
za_hbg_filtered <-
  left_join(za_hbg_filtered,
          (baz_ref %>% mutate(geog_in_zone = TRUE)), # join to list of block groups in each baz
          by = c("zone_id" = "baz", "zone_name" = "baz_name", "geoid" = "blkgrp")) %>% 
  replace_na(list(geog_in_zone = FALSE)) %>% # replace anything that didn't join with FALSE
  relocate(geog_in_zone, .after = geoid) # move to after geoid 

# remove places where the home geography is tagged as visitor
za_hbg_filtered %<>% filter(!(geog_in_zone & homework_filter == "Visitors"))
```


## States

```{r, eval = FALSE}
# read home states
za_state <- read_csv(paste0("./Data/", za, "/Home Work/", za, "_home_state_all.csv"),
                   col_types = "ccccccccdd",
                   na = c("", "N/A"))
# rename
za_state <- rename(za_state, any_of(stl_names))

# filter to all days, all hours (for highest sample size), outside MA
za_state_filtered <-
  filter(za_state,
         day_type == "0: All Days (M-Su)" &
         day_part == "0: All Day (12am-12am)" &
           state_name != "Massachusetts")
rm(za_state)
gc()

# write to RData file
save(za_state_filtered, file = paste0("./Data/", za, "_state_filtered.RData"))
```

```{r}
load(paste0("./Data/", za, "_state_filtered.RData"))
```

```{r}
za_state_filtered %<>% mutate(geog_in_zone = FALSE, .after = state_name)
```


## ZA 

```{r}
# read csv
za_all <- read_csv(paste0("./Data/", za, "/", za, "_za_all.csv"),
                  col_types = "ccccccdcccddddd",
                  na = c("", "N/A"))

# rename
za_all <- rename(za_all, any_of(stl_names))

# filter to all days, all hours (for highest sample size)
za_all_filtered <-
  filter(za_all,
         day_type == "0: All Days (M-Su)" &
         day_part == "0: All Day (12am-12am)")
rm(za_all)
gc()
```

## Residents Table

```{r}
# create clean residents table
za_residents <-
  za_all_filtered %>% 
  filter(homework_filter == "Residents") %>% 
  select(mode, homework_filter, int_type, zone_id, zone_name, day_type, day_part, stl_vol)
rm(za_all_filtered)

# create join field
# for baz over 831, za_residents has bg in both, baz_pop has them in the blkgrp but not in baz or baz_name
baz_pops_tojoin <-
  baz_pops %>% 
  transmute(zone_id = if_else(as.numeric(baz) <= 831, baz, blkgrp),
            zone_name = if_else(as.numeric(baz) <= 831, baz_name, blkgrp),
            blkgrp, bg_pop, baz_pop, bg_pct_baz)

za_residents <-
  left_join(za_residents, baz_pops_tojoin, by = c("zone_id", "zone_name")) %>% 
  transmute(mode, homework_filter, int_type, zone_id, zone_name, day_type, day_part,
            geoid = blkgrp, geog_in_zone = TRUE, stl_vol, pct_home_location = bg_pct_baz)
rm(baz_pops_tojoin)
```

## Create Final HWL Table

```{r}
# list home and work location tables to join 
hwl_list <- lst(za_hbg_filtered, za_residents, za_state_filtered)
rm(za_hbg_filtered, za_residents, za_state_filtered)
# home geography rename
home_geog_names <- c("home_geog" = "geoid", "home_geog" = "state_name")
hwl <- map(hwl_list, ~rename(.x, any_of(home_geog_names))) %>% bind_rows()
rm(hwl_list, home_geog_names)


# Calculate Percent Covered
# calculate volume by home location, with no correction for covered volume
hwl %<>% mutate(vol_home_loc_no_corr = stl_vol * pct_home_location)

# calcuate corrected volume by home location, correcting for covered volume
hwl %<>% 
  group_by(mode, homework_filter, int_type, zone_id, zone_name, day_type, day_part) %>% 
  mutate(vol_covered = sum(vol_home_loc_no_corr)) %>% 
  ungroup %>%
  mutate(pct_covered = vol_covered/stl_vol) %>% 
  mutate(vol_home_loc = vol_home_loc_no_corr/pct_covered)
```

## ZA Traveler

```{r}
# read csv
za_trav <- read_csv(paste0("./Data/", za, "/", za, "_zone_traveler_all.csv"),
                  col_types = "ccccccccccddddddddddddddddddddddddddd",
                  na = c("", "N/A"))
# rename
za_trav <- rename(za_trav, any_of(stl_names))

# filter to all days, all hours (for highest sample size)
za_trav_filtered <-
  filter(za_trav,
         day_type == "0: All Days (M-Su)" &
         day_part == "0: All Day (12am-12am)")
rm(za_trav)
gc()

# remove education & family size variables
za_trav_filtered %<>% 
  select(mode:zone_name,
         day_type:inc_more_200k,
         white:hispanic)
```


## Join to Census Data

```{r}
census_info <-
  bind_rows(
    (census_mass_info %>% select(geog_name = GEOID, pop_tot, pop_hisp_pct, pop_white_pct, pop_hisp_white_pct, pop_mino_pct, hh_li45_pct,
                                 hh_li50_pct, hh_lep_pct, hh_limveh_weak_pct)),
    (census_states_info %>% select(geog_name = NAME, pop_tot, pop_hisp_pct, pop_white_pct, pop_hisp_white_pct, pop_mino_pct, hh_li45_pct,
                                   hh_li50_pct, hh_lep_pct, hh_limveh_weak_pct))
    ) 

census_info %<>% rename_with(~str_replace(.x, "_pct", "_pct_acs"))

hwl_census_pct <- left_join(hwl, census_info, by = c("home_geog" = "geog_name"))
rm(census_mass_info, census_states_info, census_info, hwl)
gc()
```


# Math

```{r}
# convert percents to volumes
hwl_census_vol <-
  hwl_census_pct %>% 
  mutate(across(ends_with("pct_acs"), ~ vol_home_loc * .x)) %>% 
  rename_with(~str_replace(.x, "_pct_acs", "_vol_home_geog"))

# sum up to zone
hwl_volumes <-
  hwl_census_vol %>% 
  group_by(mode, homework_filter, int_type, zone_id, zone_name, day_type, day_part) %>% 
  summarize(stl_vol = max(stl_vol), # these should be the same
            check_stl_vol = sum(vol_home_loc),
            across(ends_with("vol_home_geog"), ~ sum(.x, na.rm = TRUE))) %>% 
  rename_with(~str_replace(.x, "_vol_home_geog", "_vol")) %>% 
  ungroup

rm(hwl_census_pct, hwl_census_vol)
```


# QA

```{r}
# compare % hispanic, % white, % income under 50k
hwl_comp <-
  hwl_volumes %>% 
  transmute(homework_filter, int_type, zone_id, zone_name, day_type, day_part, 
            hwl_vol = stl_vol,
            hwl_hisp_vol = pop_hisp_vol, hwl_white_vol = pop_white_vol, hwl_li50_vol = hh_li50_vol,
            hwl_hisp_pct = pop_hisp_vol/stl_vol, hwl_white_pct = pop_white_vol/stl_vol, hwl_li50_pct = hh_li50_vol/stl_vol)
za_comp <-
  za_trav_filtered %>% 
  transmute(homework_filter, int_type, zone_id, zone_name, day_type, day_part,
            za_vol = stl_vol,
            za_hisp_vol = hispanic * stl_vol, za_white_vol = white * stl_vol,
            za_li50_vol = (inc_less_20k + inc_20k_35k + inc_35k_50k) * stl_vol,
            za_hisp_pct = hispanic, za_white_pct = white, za_li50_pct = inc_less_20k + inc_20k_35k + inc_35k_50k)

comp <-
  left_join(hwl_comp, za_comp,
            by = c("homework_filter", "int_type", "zone_id", "zone_name", "day_type", "day_part"))
rm(hwl_comp, za_comp)

comp_agg <-
  comp %>% 
  group_by(int_type, zone_id, zone_name, day_type, day_part) %>% 
  summarize(across(ends_with("vol"), ~sum(.x, na.rm = TRUE))) %>% 
  ungroup

# comp %>%
#   ggplot(aes(x = hwl_hisp_vol, y = za_hisp_vol)) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic() +
#   labs(x = "Hispanic Travel Volume\n(Home & Work Locations + ACS)", y = "Hispanic Travel Volume\n(Zone Activity Analysis)") +
#   theme(axis.title.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))
# 
# comp %>%
#   ggplot(aes(x = hwl_white_vol, y = za_white_vol)) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic() +
#   labs(x = "White Travel Volume\n(Home & Work Locations + ACS)", y = "White Travel Volume\n(Zone Activity Analysis)") +
#   theme(axis.title.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))
# 
# comp %>%
#   ggplot(aes(x = hwl_li50_vol, y = za_li50_vol)) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   theme_classic() +
#   labs(x = "Low-Income Travel Volume\n(Home & Work Locations + ACS)", y = "Low-Income Travel Volume\n(Zone Activity Analysis)") +
#   theme(axis.title.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))


comp_agg %>% 
  ggplot(aes(x = hwl_hisp_vol, y = za_hisp_vol)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "Hispanic Travel Volume\n(Home & Work Locations + ACS)", y = "Hispanic Travel Volume\n(Zone Activity Analysis)") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))

comp_agg %>% 
  ggplot(aes(x = hwl_white_vol, y = za_white_vol)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "White Travel Volume\n(Home & Work Locations + ACS)", y = "White Travel Volume\n(Zone Activity Analysis)") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))

comp_agg %>% 
  ggplot(aes(x = hwl_li50_vol, y = za_li50_vol)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  labs(x = "Low-Income Travel Volume\n(Home & Work Locations + ACS)", y = "Low-Income Travel Volume\n(Zone Activity Analysis)") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))
# all changes make directional sense, given that StreetLight ZA uses 2010 census, and this analysis uses 2019 ACS.

rm(comp, comp_agg)
```


# Final Output (Zone/IntType Level)

```{r}
hwl_za <-
  left_join(
  (hwl_volumes %>% 
    select(mode:day_part,
           hwl_vol = stl_vol, hwl_white_vol = pop_white_vol, hwl_hisp_white_vol = pop_hisp_white_vol, hwl_mino_vol = pop_mino_vol,
           hwl_li45_vol = hh_li45_vol, hwl_li50_vol = hh_li50_vol, hwl_lep_vol = hh_lep_vol, hwl_limveh_weak_vol = hh_limveh_weak_vol)),
  (za_trav_filtered %>% 
    transmute(mode, homework_filter, int_type, zone_id, zone_name, day_type, day_part,
              za_vol = stl_vol,
              za_white_vol = white * stl_vol, za_li50_vol = (inc_less_20k + inc_20k_35k + inc_35k_50k) * stl_vol)),
  by = c("mode", "homework_filter", "int_type", "zone_id", "zone_name", "day_type", "day_part")
)
rm(hwl_volumes, za_trav_filtered)

# int type, zone id, zone name, day type, day part
# stl vol
# percent of white travel that is made by white hispanic people
# percent of total travel made by white hispanic people
# percent of total travel made by non- white hispanic people
# percent of travel made by low-income (45 & 50k) people
# percent of travel made by people with limited access to a vehicle (exclusive of 0.5)
# percent of travel made by people with limited english proficiency

hwl_za_agg <-
  hwl_za %>% 
  group_by(mode, int_type, zone_id, zone_name, day_type, day_part) %>% 
  summarize(across(ends_with("vol"), ~ sum(.x, na.rm = TRUE))) %>% 
  ungroup

demo_corrections_output <-
  hwl_za_agg %>% 
  mutate(hisp_white_of_tot_pct = hwl_hisp_white_vol/hwl_vol,
         hisp_white_of_white_pct = hwl_hisp_white_vol/hwl_white_vol,
         mino_pct = hwl_mino_vol/hwl_vol,
         li45_pct = hwl_li45_vol/hwl_vol,
         li50_pct = hwl_li50_vol/hwl_vol,
         limveh_weak_pct = hwl_limveh_weak_vol/hwl_vol,
         lep_vol = hwl_lep_vol/hwl_vol)
save(demo_corrections_output, file = paste0("./Data/Demo_Corrections_Output_", za, ".RData"))
write_csv(demo_corrections_output, paste0("./Data/Demo_Corrections_Output_", za, ".csv"))
rm(hwl_za, hwl_za_agg)
gc()
```


# Apply to O-D

```{r}
load(paste0("./Data/Demo_Corrections_Output_", za, ".RData"))
```


## Read Bike Analysis

```{r, eval = FALSE}
bike_trav <- read_csv(paste0("./Data/", bike, "/", bike, "_od_traveler_bike.csv"),
                      col_types = "ccccdccccdcccdddddddddddddddddddddddddddddd",
                      na = c("", "N/A"))
bike_trav %<>% rename(any_of(stl_names))

bike_trav %<>% 
  transmute(orig_zone_id, orig_zone_name, dest_zone_id, dest_zone_name,
            day_type, day_part,
            b_stl_inc_less_50k_pct = inc_less_20k + inc_20k_35k + inc_35k_50k,
            b_stl_white_pct = white,
            b_purpose_hbw_pct = purpose_hbw,
            b_purpose_nhb_pct = purpose_nhb,
            b_purpose_hbo_pct = purpose_hbo)


bike_trip <- read_csv(paste0("./Data/", bike, "/", bike, "_od_trip_bike.csv"),
                      col_types = "ccccdccccdcccddddddddddddddddddddddddddddddddddddd",
                      na = c("", "N/A"))
bike_trip %<>% rename(any_of(stl_names))
bike_trip %<>% 
  transmute(orig_zone_id, orig_zone_name, dest_zone_id, dest_zone_name,
            day_type, day_part,
            b_ind = stl_ind,
            b_short_ind = stl_ind * trip_less_1mi,
            b_long_ind = stl_ind * trip_more_1mi)
  

save(bike_trav, bike_trip, file = paste0("./Data/", bike, "_cleaned.RData"))
```

```{r}
load(paste0("./Data/", bike, "_cleaned.RData"))
```


## Read Ped Analysis


```{r, eval = FALSE}
ped_trav <- read_csv(paste0("./Data/", ped, "/", ped, "_od_traveler_ped.csv"),
                      col_types = "ccccdccccdcccdddddddddddddddddddddddddddddd",
                      na = c("", "N/A"))
ped_trav %<>% rename(any_of(stl_names))

ped_trav %<>% 
  transmute(orig_zone_id, orig_zone_name, dest_zone_id, dest_zone_name,
            day_type, day_part,
            p_stl_inc_less_50k_pct = inc_less_20k + inc_20k_35k + inc_35k_50k,
            p_stl_white_pct = white,
            p_purpose_hbw_pct = purpose_hbw,
            p_purpose_nhb_pct = purpose_nhb,
            p_purpose_hbo_pct = purpose_hbo)

ped_trip <- read_csv(paste0("./Data/", ped, "/", ped, "_od_trip_ped.csv"),
                      col_types = "ccccdccccdcccdddddddddddddddddddddddddddddddddd",
                      na = c("", "N/A"))
ped_trip %<>% rename(any_of(stl_names))
ped_trip %<>% 
  transmute(orig_zone_id, orig_zone_name, dest_zone_id, dest_zone_name,
            day_type, day_part,
            p_ind = stl_ind,
            p_short_ind = stl_ind * trip_less_1mi,
            p_long_ind = stl_ind * trip_more_1mi)

save(ped_trav, ped_trip, file = paste0("./Data/", ped, "_cleaned.RData"))
```

```{r}
load(paste0("./Data/", ped, "_cleaned.RData"))
```

## Read Vehicle Analysis

```{r, eval = FALSE}
veh_trav <- read_csv(paste0("./Data/", veh, "/", veh, "_od_traveler_all.csv"),
                      col_types = "ccccdccccdcccdddddddddddddddddddddddddddddd",
                      na = c("", "N/A"))
veh_trav %<>% rename(any_of(stl_names))
veh_trav %<>% 
  transmute(orig_zone_id, orig_zone_name, dest_zone_id, dest_zone_name,
            day_type, day_part,
            v_stl_inc_less_50k_pct = inc_less_20k + inc_20k_35k + inc_35k_50k,
            v_stl_white_pct = white,
            v_purpose_hbw_pct = purpose_hbw,
            v_purpose_nhb_pct = purpose_nhb,
            v_purpose_hbo_pct = purpose_hbo)
gc()

veh_trip <- read_csv(paste0("./Data/", veh, "/", veh, "_od_trip_all.csv"),
                      col_types = veh_trips_col_types,
                      na = c("", "N/A")
                     )
veh_trip %<>% 
  rename(any_of(stl_names)) 
veh_trip %<>% 
  transmute(orig_zone_id, orig_zone_name, dest_zone_id, dest_zone_name,
            day_type, day_part,
            v_vol = stl_vol,
            v_short_vol = stl_vol * trip_less_1mi,
            v_long_vol = stl_vol * trip_more_1mi)
gc()

save(veh_trav, veh_trip, file = paste0("./Data/", veh, "_cleaned.RData"))
```


```{r}
load(paste0("./Data/", veh, "_cleaned.RData"))
```


## Create Combined Bike/Ped/Vehicle

```{r}
join_fields <- c("orig_zone_id", "orig_zone_name", "dest_zone_id", "dest_zone_name", "day_type", "day_part")

bike_comb <-
  left_join(bike_trav, bike_trip,
            by = join_fields)
bike_comb %<>% 
  mutate(across(ends_with("_pct"), ~.x * b_ind)) %>% 
  rename_with(~str_replace(.x, "_pct", "_ind"))
rm(bike_trav, bike_trip)

ped_comb <-
  left_join(ped_trav, ped_trip,
            by = join_fields)
ped_comb %<>% 
  mutate(across(ends_with("_pct"), ~.x * p_ind)) %>% 
  rename_with(~str_replace(.x, "_pct", "_ind"))
rm(ped_trav, ped_trip)

veh_comb <-
  left_join(veh_trav, veh_trip,
            by = join_fields)
veh_comb %<>% 
  mutate(across(ends_with("_pct"), ~.x * v_vol)) %>% 
  rename_with(~str_replace(.x, "_pct", "_vol"))
rm(veh_trav, veh_trip)
gc()
```


## Merge Files

```{r}
bike_ped_comb <- full_join(bike_comb, ped_comb, by = join_fields)
rm(bike_comb, ped_comb)

bike_ped_veh_comb <- full_join(bike_ped_comb, veh_comb, by = join_fields)
rm(bike_ped_comb, veh_comb)
gc()

rm(join_fields)
```


## Math

```{r}
# replace NAs
bike_ped_veh_comb %<>% 
  mutate(across(where(is.double), ~ replace_na(., 0)))

bike_ped_veh_vols <-
  bike_ped_veh_comb %>% 
  transmute(orig_zone_id, orig_zone_name, dest_zone_id, dest_zone_name, 
            day_type, day_part,
            stl_vol = b_ind + p_ind + v_vol,
            stl_inc_less_50k_vol = b_stl_inc_less_50k_ind + p_stl_inc_less_50k_ind + v_stl_inc_less_50k_vol,
            stl_white_vol = b_stl_white_ind + p_stl_white_ind + v_stl_white_vol,
            purpose_hbw_vol = b_purpose_hbw_ind + p_purpose_hbw_ind + v_purpose_hbw_vol,
            purpose_nhb_vol = b_purpose_nhb_ind + p_purpose_nhb_ind + v_purpose_nhb_vol,
            purpose_hbo_vol = b_purpose_hbo_ind + p_purpose_hbo_ind + v_purpose_hbo_vol,
            short_vol = b_short_ind + p_short_ind + v_short_vol,
            long_vol = b_long_ind + p_long_ind + v_long_vol)
rm(bike_ped_veh_comb)
gc()
```
## Apply Demo Corrections on Start and End

```{r}
dc_orig <-
  demo_corrections_output %>% 
  filter(int_type == "Trip Start") %>% 
  select(zone_id, zone_name, hwl_vol, hisp_white_of_white_pct, li50_pct, limveh_weak_pct, mino_pct) %>% 
  rename_with(~str_c("orig_", .x))
dc_dest <-
  demo_corrections_output %>%
  filter(int_type == "Trip End") %>%
  select(zone_id, zone_name, hwl_vol, hisp_white_of_white_pct, li50_pct, limveh_weak_pct, mino_pct) %>% 
  rename_with(~str_c("dest_", .x))

bike_ped_veh_vols_joined <-
  bike_ped_veh_vols %>% 
  left_join(dc_orig, by = c("orig_zone_id", "orig_zone_name")) %>% 
  left_join(dc_dest, by = c("dest_zone_id", "dest_zone_name")) 


bike_ped_veh_vols_joined %<>% 
  mutate(hisp_white_of_white_pct = ((orig_hwl_vol * orig_hisp_white_of_white_pct) + (dest_hwl_vol * dest_hisp_white_of_white_pct))/
                                      (orig_hwl_vol + dest_hwl_vol),
         mino_pct = ((orig_hwl_vol * orig_mino_pct) + (dest_hwl_vol * dest_mino_pct))/(orig_hwl_vol + dest_hwl_vol),
         li50_pct = ((orig_hwl_vol * orig_li50_pct) + (dest_hwl_vol * dest_li50_pct))/(orig_hwl_vol + dest_hwl_vol),
         limveh_pct = ((orig_hwl_vol * orig_limveh_weak_pct) + (dest_hwl_vol * dest_limveh_weak_pct))/(orig_hwl_vol + dest_hwl_vol))

od_output <-
  bike_ped_veh_vols_joined %>% 
  transmute(orig_zone_id, orig_zone_name,
            dest_zone_id, dest_zone_name,
            day_type, day_part,
            stl_vol,
            mino_vol_hwl = stl_vol * mino_pct,
            mino_vol_stlcor = stl_vol - (stl_white_vol * (1 - hisp_white_of_white_pct)),
            li50_vol_hwl = stl_vol * li50_pct,
            li50_vol_stl = stl_inc_less_50k_vol,
            limveh_vol = stl_vol * limveh_pct,
            purpose_hbw_vol,
            purpose_hbo_vol,
            purpose_nhb_vol,
            short_vol,
            long_vol)

od_output %<>% 
  transmute(orig_zone_id, orig_zone_name,
            dest_zone_id, dest_zone_name,
            day_type, day_part,
            stl_vol,
            mino_vol = if_else(is.na(mino_vol_hwl), mino_vol_stlcor, mino_vol_hwl),
            li50_vol = if_else(is.na(li50_vol_hwl), li50_vol_stl, li50_vol_hwl),
            limveh_vol,
            purpose_hbw_vol,
            purpose_hbo_vol,
            purpose_nhb_vol,
            short_vol,
            long_vol)

od_output %<>% arrange(orig_zone_id, dest_zone_id, day_type, day_part)
```

# OD QA

```{r, eval = FALSE}
od_output_testing <-
  bike_ped_veh_vols_joined %>% 
  transmute(orig_zone_id, orig_zone_name,
            dest_zone_id, dest_zone_name,
            day_type, day_part,
            stl_vol,
            mino_vol_hwl = stl_vol * mino_pct,
            mino_vol_stlcor = stl_vol - (stl_white_vol * (1 - hisp_white_of_white_pct)),
            li50_vol_hwl = stl_vol * li50_pct,
            li50_vol_stl = stl_inc_less_50k_vol,
            limveh_vol = stl_vol * limveh_pct,
            purpose_hbw_vol,
            purpose_hbo_vol,
            purpose_nhb_vol,
            short_vol,
            long_vol)

# demo_corrections_output %>% filter(zone_id == "250173164006")
od_malden_everett <- 
  od_output_testing %>%  
  filter(str_detect(orig_zone_name, "Malden") | str_detect(orig_zone_name, "Everett") |
           str_detect(dest_zone_name, "Malden") | str_detect(dest_zone_name, "Everett"))


od_malden_everett %>%   
  ggplot(aes(x = mino_vol_hwl, y = mino_vol_stlcor)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() + xlim(0,5000) + ylim(0,5000) +
  labs(x = "Minority Travel Volume\n(Home & Work Locations + ACS)", y = "Minority Travel Volume\n(Corrected StL OD)") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))

od_malden_everett %>% 
  ggplot(aes(x = li50_vol_hwl, y = li50_vol_stl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  # xlim(0,5000) + ylim(0,5000) +
  labs(x = "Low-Income Travel Volume\n(Home & Work Locations + ACS)", y = "Low-Income Travel Volume\n(StreetLight OD)") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))


od_output_testing %>% slice_sample(prop = 0.05) %>%    
  ggplot(aes(x = mino_vol_hwl, y = mino_vol_stlcor)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() + 
  # xlim(0,5000) + ylim(0,5000) +
  labs(x = "Minority Travel Volume\n(Home & Work Locations + ACS)", y = "Minority Travel Volume\n(Corrected StL OD)") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))

od_output_testing %>% slice_sample(prop = 0.05) %>%    
  ggplot(aes(x = li50_vol_hwl, y = li50_vol_stl)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_classic() +
  # xlim(0,5000) + ylim(0,5000) +
  labs(x = "Low-Income Travel Volume\n(Home & Work Locations + ACS)", y = "Low-Income Travel Volume\n(StreetLight OD)") +
  theme(axis.title.y = element_text(angle = 0, vjust = 0.5, hjust = 0.5))

od_output_testing %>% slice_sample(prop = 0.05) %>% 
  filter(is.na(li50_vol_hwl) | is.na(li50_vol_stl))
```

# Output OD

```{r}
save(od_output, file = paste0("./Data/", od_output_name, ".RData"))
write_csv(od_output, paste0("./Data/", od_output_name, ".csv"))
```

# Clean Up

```{r}
rm(baz_pops, baz_ref, bike_ped_veh_vols, bike_ped_veh_vols_joined, dc_dest, dc_orig)
rm(bike, ped, veh, za, user, stl_names, od_output_name, veh_trips_col_types)
gc()
```


